---
title: "Statistica Learning"
author: "Arielle"
date: "11/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library (tidyverse)
library (viridis)
library (gridExtra)
library(modelr)
library(mgcv)
library(patchwork)
library(glmnet)


knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))

set.seed(11)
```

Lets start with LASSO

```{r}
bwt_df = 
  read_csv("./data/extra_topic_data/birthweight.csv") %>% 
  janitor::clean_names() %>%
  mutate(
    babysex = as.factor(babysex),
    babysex = fct_recode(babysex, "male" = "1", "female" = "2"),
    frace = as.factor(frace),
    frace = fct_recode(frace, "white" = "1", "black" = "2", "asian" = "3", 
                       "puerto rican" = "4", "other" = "8"),
    malform = as.logical(malform),
    mrace = as.factor(mrace),
    mrace = fct_recode(mrace, "white" = "1", "black" = "2", "asian" = "3", 
                       "puerto rican" = "4")) %>% 
  sample_n(200)
```

To use the lasso, we will use glmnet

```{r}
y = bwt_df$bwt

x = model.matrix(bwt ~ ., bwt_df)[,-1]
```

```{r}
lasso_fit = glmnet(x, y)
```

Make sure you set those lambdas to make sure its the same

```{r}
lambda = 10^(seq(3, -2, -0.1))

lasso_fit =
  glmnet(x, y, lambda = lambda)

lasso_cv =
  cv.glmnet(x, y, lambda = lambda)

lambda_opt = lasso_cv$lambda.min
```

Still going to use the broom packeage on the lasso fit because it will look better 

```{r}
broom::tidy(lasso_fit) %>% 
  select(term, lambda, estimate) %>% 
  complete(term, lambda, fill = list(estimate = 0) ) 
```

```{r}
broom::tidy(lasso_fit) %>% 
  select(term, lambda, estimate) %>% 
  complete(term, lambda, fill = list(estimate = 0) ) %>% 
  filter(term != "(Intercept)") %>% 
  ggplot(aes(x = log(lambda, 10), y = estimate, group = term, color = term)) + 
  geom_path() + 
  geom_vline(xintercept = log(lambda_opt, 10), color = "blue", size = 1.2) +
  theme(legend.position = "none")
```

This is the cross validation part 

```{r}
broom::tidy(lasso_cv) %>% 
  ggplot(aes(x = log(lambda, 10), y = estimate)) + 
  geom_point() 
```

The blue line in the first model is the smallest lamnda will be for the set of predictions thats are present from the second graph. 


### CLustering 

```{r}
poke_df = 
  read_csv("./data/extra_topic_data/pokemon.csv") %>% 
  janitor::clean_names() %>% 
  select(hp, speed)
```

```{r}
poke_df %>% 
  ggplot(aes(x = hp, y = speed)) + 
  geom_point()
```


We clustersing now 

```{r}
kmeans_fit =
  kmeans(x = poke_df, centers = 3)
```


We processing and plotting 

```{r}
poke_df =
  broom::augment(kmeans_fit, poke_df)

poke_df %>% 
  ggplot(aes(x = hp, y = speed, color = .cluster)) +
  geom_point()
```

### Clustering Trajectories 

```{r}
traj_data = 
  read_csv("./data/extra_topic_data/trajectories.csv")

traj_data %>% 
  ggplot(aes(x = week, y = value, group = subj)) + 
  geom_point() + 
  geom_path()
```

Now we need to get intercepts and slopes for eveyone 
```{r}
int_slope_df = 
  traj_data %>% 
  nest(data = week:value) %>% 
  mutate(
    models = map(data, ~lm(value ~ week, data = .x)),
    result = map(models, broom::tidy)
  ) %>% 
  select(-data, -models) %>% 
  unnest(result) %>% 
  select(subj, term, estimate) %>% 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) %>% 
  rename(int = "(Intercept)", slope = week)
```

Try to kmeans this but first we gonna plot the intercept and slope

```{r}
int_slope_df %>% 
  ggplot(aes(x = int, y = slope)) + 
  geom_point()
```

Have to do some processing, because it will treat the sunject as something to be clustered but thats not the case

```{r}
km_fit = 
  kmeans(
    x = int_slope_df %>% 
      select(-subj) %>% 
      scale,  centers = 2)

int_slope_df =
  broom::augment(km_fit, int_slope_df)
```

We clustered this

```{r}
int_slope_df %>% 
  ggplot(aes(x = int, y = slope, color = .cluster)) +
  geom_point()
```

We joined the two together

```{r}
left_join(traj_data, int_slope_df) %>% 
  ggplot(aes(x = week, y = value, group = subj, color = .cluster)) + 
  geom_point() + 
  geom_path() 
```

